---
title: "Covid91 vaccine study Final2024"
author: "Daniel Esteves"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "Summer 2024" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="finalSumr2024.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

## Males

$H_0$: The Covid-921 vaccine is not effective in males.

$H_A$: The Covid-921 vaccine is effective in males.

### methods

Treatment - categorical), Male - categorical
Helper File: CAT~CAT

### Descriptive Results

#### Graphical Results


```{r}
males=filter(FinalData,FinalData$sex=="M")
barchartGC(~treatment + infected,data=males)
barchartGC(~infected + treatment,data=males,type="percent")

```

This graph shows that the males who got the placebo treatment were more infected than the males that were vaccinated. However, about 50% of males with placebos and about 50% of males with the vaccine were not infected. 

#### Numerical Results

```{r}
table1 <- xtabs(~treatment + infected,data=males)
rowPerc(table1)
colPerc(table1)
```
This shows that about 74% of males who took the placebo got COVID-19, whereas about 26% of males who got the vaccine got COVID-19.

#### Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```


The analysis shows a difference in COVID-19 infection rates between the placebo and vaccine groups. Both the chi-square test and Fisher’s Exact Test yielded a p-value of 0.00000000022, indicating a strong reduction in the risk of contracting COVID-19 for those vaccinated. Specifically, individuals receiving the placebo are about 2.82 times more likely to contract COVID-19 compared to those who received the vaccine. The 95% confidence interval further supports the vaccine’s high effectiveness.

## Females

$H_0$: The Covid921 vaccine is not effective in females

$H_A$: The Covid921 vaccine is effective in females

### Methods


Treatment-categorical), Female-categorical
Helper File: CAT ~ CAT

### Descriptive Results

#### Graphical Results

```{r}
females=filter(FinalData,FinalData$sex=="F")
barchartGC(~treatment + infected,data=females)
barchartGC(~infected + treatment,data=females,type="percent")

```

This graph shows that the females who got the placebo were more infected than the females that were vaccinated. However, about 50% of females with placebos and about 51% of females with the vaccine were not infected.  

### Numerical Results

```{r}
table1 <- xtabs(~treatment + infected,data=males)
rowPerc(table1)
colPerc(table1)
```

This reveals that the COVID-19 vaccine significantly decreases the likelihood of catching the virus compared to the placebo. Specifically, those who receive the vaccine have approximately 46.9% lower odds of getting COVID-19 compared to those who receive the placebo. The relative risk reduction is about 45.5%, meaning that the vaccine reduces the risk of infection by about 45.5% relative to the placebo.

#### Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The chi-square test result of 0.00000001914 and Fisher’s Exact Test result of 0.00000001246 both confirm the statistical significance of the difference in infection rates between the vaccine and placebo groups. The odds ratio of 1.8562 indicates that the likelihood of contracting COVID-19 with the placebo is about 1.86 times higher than with the vaccine. The 95% confidence interval, ranging from 1.49 to 2.32, further supports the effectiveness of the vaccine.  


## LGBTQ

$H_0$: The Covid-921 vaccine is not effective in preventing COVID-19 in the LGBTQ community.

$H_A$: The Covid-921 vaccine is effective in preventing COVID-19 in the LGBTQ community.

### Methods

Variables: LGBTQ-categorical, Treatment-categorical
Helper File: CAT~CAT



## Descriptive Results

### Graphical Results

```{r}
LGBTQData <- filter(FinalData, LGBTQ == "gay")
barchartGC(~infected + treatment, data=LGBTQData)
barchartGC(~infected + treatment, data=LGBTQData, type="percent")
```


This graph shows that the people in the LGBTQ community who got the placebo were more infected than those who were vaccinated. However, about 45% of those with placebos and about 49% of those with the vaccine were not infected.

### Numerical Descriptive Results

```{r}
table3 <- xtabs(~infected + treatment, data=LGBTQData)
rowPerc(table3)
colPerc(table3)

```

The results show that the COVID-19 vaccine effectively reduces the risk of contracting the virus. People who received the vaccine have about a 24% chance of getting COVID-19 compared to those who received a placebo. This means the vaccine reduces the infection risk by about 44.8% compared to the placebo. Overall, the vaccine offers a significant protective effect with a relative risk reduction of approximately 55.2%.


## Infrenital Results

```{r}
chisq.test(table3)
chisqtestGC(table3)
fisher.test(table3)

```



The COVID-19 vaccine greatly reduces the risk of getting the virus. People with the vaccine have only 24% of the chance of infection compared to those with a placebo. This means the vaccine cuts the infection risk by about 44.8%. Overall, the vaccine reduces the risk by about 55.2%.


## Druggies

$H_0$: The Covid-921 vaccine is not effective in preventing COVID-19 in drug users.

$H_A$: The Covid-921 vaccine is effective in preventing COVID-19 in drug users.


### Methods

Drug User-categorical, Treatment-categorical
Helper File: CAT~CAT

## Descriptive Results

### Graphical Results

```{r}
DruggiesData <- filter(FinalData, DrugUser == "yes")
barchartGC(~infected + treatment, data=DruggiesData)
barchartGC(~infected + treatment, data=DruggiesData, type="percent")

```


This graph shows that the drug users who got the placebo were less likely to get COVID-19 than those who got the vaccine, with both groups having about 50% remaining uninfected.

### Numerical Descriptive Results

```{r}
table4 <- xtabs(~infected + treatment, data=DruggiesData)
rowPerc(table4)
colPerc(table4)

```

The data reveals a 3.28% difference in COVID-19 infection rates between the groups. Specifically, 3.61% of those who received the vaccine contracted COVID-19, compared to 0.33% of those who received the placebo. Among those who remained healthy, there was a difference of only about 2.61%. This unusual finding suggests the vaccine may increase the risk of infection compared to the placebo, which could indicate an anomaly in the data or a specific aspect of the study that warrants further investigation. Typically, one would expect a vaccine to reduce, rather than increase, the risk of COVID-19.  

## Infrenital Results

```{r}
chisq.test(table4)
chisqtestGC(table4)
fisher.test(table4)

```

The p-value from the Chi-square Test is 0.0000000001414, suggesting that the vaccine does not effectively prevent COVID-19 among drug users. Similarly, Fisher's Exact Test yields a p-value of 0.000000000003261, indicating a minor difference in COVID-19 rates among drug users. We do not reject the null hypothesis because drug users can still contract COVID-19 regardless of whether they receive the vaccine or not. 





# Overall Results and Conclusions