---
title: "Covid91 vaccine study Final2024"
author: "Daniel Esteves"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "Summer 2024" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="finalSumr2024.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

## Males

$H_0$: The Covid-921 vaccine is not effective in males.

$H_A$: The Covid-921 vaccine is effective in males.

### methods

Treatment - categorical), Male - categorical
Helper File: CAT~CAT

### Descriptive Results

#### Graphical Results


```{r}
males=filter(FinalData,FinalData$sex=="M")
barchartGC(~treatment + infected,data=males)
barchartGC(~infected + treatment,data=males,type="percent")

```

This graph shows that the males who got the placebo treatment were more infected than the males that were vaccinated. However, about 50% of males with placebos and about 50% of males with the vaccine were not infected. 

#### Numerical Results

```{r}
table1 <- xtabs(~treatment + infected,data=males)
rowPerc(table1)
colPerc(table1)
```
This shows that about 74% of males who took the placebo got COVID-19, whereas about 26% of males who got the vaccine got COVID-19.

#### Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```


The analysis shows a difference in COVID-19 infection rates between the placebo and vaccine groups. Both the chi-square test and Fisher’s Exact Test yielded a p-value of 0.00000000022, indicating a strong reduction in the risk of contracting COVID-19 for those vaccinated. Specifically, individuals receiving the placebo are about 2.82 times more likely to contract COVID-19 compared to those who received the vaccine. The 95% confidence interval further supports the vaccine’s high effectiveness.

## Females

$H_0$: The Covid921 vaccine is not effective in females

$H_A$: The Covid921 vaccine is effective in females

### Methods


Treatment-categorical), Female-categorical
Helper File: CAT ~ CAT

### Descriptive Results

#### Graphical Results

```{r}
females=filter(FinalData,FinalData$sex=="F")
barchartGC(~treatment + infected,data=females)
barchartGC(~infected + treatment,data=females,type="percent")

```

This graph shows that the females who got the placebo were more infected than the females that were vaccinated. However, about 50% of females with placebos and about 51% of females with the vaccine were not infected.  

### Numerical Results

```{r}
table1 <- xtabs(~treatment + infected,data=males)
rowPerc(table1)
colPerc(table1)
```

This reveals that the COVID-19 vaccine significantly decreases the likelihood of catching the virus compared to the placebo. Specifically, those who receive the vaccine have approximately 46.9% lower odds of getting COVID-19 compared to those who receive the placebo. The relative risk reduction is about 45.5%, meaning that the vaccine reduces the risk of infection by about 45.5% relative to the placebo.


## LGBTQ



## Druggies


# Overall Results and Conclusions