---
title: "Covid91 vaccine study Final2024"
author: "Daniel Esteves"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "Summer 2024" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="finalSumr2024.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

## Males

$H_0$: The Covid-921 vaccine is not effective in males.

$H_A$: The Covid-921 vaccine is effective in males.

### methods

Treatment - categorical), Male - categorical
Helper File: CAT~CAT

### Descriptive Results

#### Graphical Results


```{r}
males=filter(FinalData,FinalData$sex=="M")
barchartGC(~treatment + infected,data=males)
barchartGC(~infected + treatment,data=males,type="percent")

```

This graph shows that the males who got the placebo treatment were more infected than the males that were vaccinated. However, about 50% of males with placebos and about 50% of males with the vaccine were not infected. 

#### Numerical Results

```{r}
table1 <- xtabs(~treatment + infected,data=males)
rowPerc(table1)
colPerc(table1)
```
This shows that about 74% of males who took the placebo got COVID-19, whereas about 26% of males who got the vaccine got COVID-19.

## Females


## LGBTQ



## Druggies


# Overall Results and Conclusions